### 2016-09-29 doxygen/latex/annotated.tex

\section{Class List}
Here are the classes, structs, unions and interfaces with brief descriptions\+:\begin{DoxyCompactList}
\item\contentsline{section}{\hyperlink{class_blade__surface}{Blade\+\_\+surface} }{\pageref{class_blade__surface}}{}
\item\contentsline{section}{\hyperlink{class_bounding__box}{Bounding\+\_\+box} }{\pageref{class_bounding__box}}{}
\item\contentsline{section}{\hyperlink{class_point3_d}{Point3D} }{\pageref{class_point3_d}}{}
\item\contentsline{section}{\hyperlink{class_ray3_d}{Ray3D} }{\pageref{class_ray3_d}}{}
\item\contentsline{section}{\hyperlink{class_receiver}{Receiver} }{\pageref{class_receiver}}{}
\item\contentsline{section}{\hyperlink{class_rib}{Rib} }{\pageref{class_rib}}{}
\item\contentsline{section}{\hyperlink{class_rotor}{Rotor} }{\pageref{class_rotor}}{}
\item\contentsline{section}{\hyperlink{class_scene}{Scene} }{\pageref{class_scene}}{}
\item\contentsline{section}{\hyperlink{class_sphere}{Sphere} }{\pageref{class_sphere}}{}
\item\contentsline{section}{\hyperlink{class_transmitter}{Transmitter} }{\pageref{class_transmitter}}{}
\item\contentsline{section}{\hyperlink{class_triangle}{Triangle} }{\pageref{class_triangle}}{}
\item\contentsline{section}{\hyperlink{class_vector3_d}{Vector3D} }{\pageref{class_vector3_d}}{}
\end{DoxyCompactList}