### 2016-09-29 doxygen/latex/class_bounding__box.tex

\hypertarget{class_bounding__box}{}\section{Bounding\+\_\+box Class Reference}
\label{class_bounding__box}\index{Bounding\+\_\+box@{Bounding\+\_\+box}}

Collaboration diagram for Bounding\+\_\+box\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=180pt]{doxygen/latex/class_bounding__box__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_bounding__box_adf45c090850ee7b48b59f8e586450ef3}{}\label{class_bounding__box_adf45c090850ee7b48b59f8e586450ef3} 
{\bfseries Bounding\+\_\+box} (std\+::vector$<$ \hyperlink{class_point3_d}{Point3D} $>$ \&points)
\item 
\hypertarget{class_bounding__box_a29a6d043ac2de9fb152f39c1139e42df}{}\label{class_bounding__box_a29a6d043ac2de9fb152f39c1139e42df} 
bool {\bfseries hit} (const \hyperlink{class_ray3_d}{Ray3D} \&ray, \hyperlink{class_point3_d}{Point3D} \&hit\+\_\+point) const
\item 
\hypertarget{class_bounding__box_af2a1aa9cee9c82c2dd2df602cca9f42f}{}\label{class_bounding__box_af2a1aa9cee9c82c2dd2df602cca9f42f} 
bool {\bfseries hit} (const \hyperlink{class_ray3_d}{Ray3D} \&ray) const
\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_bounding__box_add720cf740129c859bddbbed01727c48}{}\label{class_bounding__box_add720cf740129c859bddbbed01727c48} 
\hyperlink{class_point3_d}{Point3D} {\bfseries min\+\_\+point}
\item 
\hypertarget{class_bounding__box_a1ee058e226dbf771065c67d62cac8c66}{}\label{class_bounding__box_a1ee058e226dbf771065c67d62cac8c66} 
\hyperlink{class_point3_d}{Point3D} {\bfseries max\+\_\+point}
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Bounding\+\_\+box.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Bounding\+\_\+box.\+cpp\end{DoxyCompactItemize}