### 2016-09-29 doxygen/latex/class_point3_d.tex

\hypertarget{class_point3_d}{}\section{Point3D Class Reference}
\label{class_point3_d}\index{Point3D@{Point3D}}

Collaboration diagram for Point3D\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=161pt]{doxygen/latex/class_point3_d__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_point3_d_a72799f11be08ba98bb37cfcf2fa0bb1f}{}\label{class_point3_d_a72799f11be08ba98bb37cfcf2fa0bb1f} 
{\bfseries Point3D} (const double x, const double y, const double z)
\item 
\hypertarget{class_point3_d_af963fbdffd449c8fc2d0631992a996e1}{}\label{class_point3_d_af963fbdffd449c8fc2d0631992a996e1} 
{\bfseries Point3D} (const double y, const double z)
\item 
\hypertarget{class_point3_d_a7ac9d9c06fe0497d868537d2f6ad9454}{}\label{class_point3_d_a7ac9d9c06fe0497d868537d2f6ad9454} 
\hyperlink{class_point3_d}{Point3D} {\bfseries operator-\/} () const
\item 
\hypertarget{class_point3_d_a6cb0fe83bc148a42fb2cf6f9aa907242}{}\label{class_point3_d_a6cb0fe83bc148a42fb2cf6f9aa907242} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries operator-\/} (const \hyperlink{class_point3_d}{Point3D} \&p) const
\item 
\hypertarget{class_point3_d_a2ab1c2f322e1845e94414fab67139667}{}\label{class_point3_d_a2ab1c2f322e1845e94414fab67139667} 
\hyperlink{class_point3_d}{Point3D} {\bfseries operator+} (const \hyperlink{class_vector3_d}{Vector3D} \&v) const
\item 
\hypertarget{class_point3_d_a376bb34529614f69a1d79b9633b6ab03}{}\label{class_point3_d_a376bb34529614f69a1d79b9633b6ab03} 
\hyperlink{class_point3_d}{Point3D} {\bfseries operator-\/} (const \hyperlink{class_vector3_d}{Vector3D} \&v) const
\item 
\hypertarget{class_point3_d_a30f07cec680e5e619289ba0c36c00d34}{}\label{class_point3_d_a30f07cec680e5e619289ba0c36c00d34} 
\hyperlink{class_point3_d}{Point3D} {\bfseries operator$\ast$} (const double a) const
\item 
\hypertarget{class_point3_d_a80016919138ae619c4ae603e0c963460}{}\label{class_point3_d_a80016919138ae619c4ae603e0c963460} 
bool {\bfseries operator==} (const \hyperlink{class_point3_d}{Point3D} \&p) const
\item 
\hypertarget{class_point3_d_a3db617345c47ecf25cee2d8bb08712dc}{}\label{class_point3_d_a3db617345c47ecf25cee2d8bb08712dc} 
double {\bfseries distance} (const \hyperlink{class_point3_d}{Point3D} \&p) const
\item 
\hypertarget{class_point3_d_aa2c65a74cd2605f2f46cd7fa88ed537c}{}\label{class_point3_d_aa2c65a74cd2605f2f46cd7fa88ed537c} 
double {\bfseries x} () const
\item 
\hypertarget{class_point3_d_aeeea1fe4bd6b56727b50c569b71f8207}{}\label{class_point3_d_aeeea1fe4bd6b56727b50c569b71f8207} 
double {\bfseries y} () const
\item 
\hypertarget{class_point3_d_ad663c810a730a64d1e000e366fc366ea}{}\label{class_point3_d_ad663c810a730a64d1e000e366fc366ea} 
double {\bfseries z} () const
\item 
\hypertarget{class_point3_d_a02de7d72811aef8bfc6660dd99267682}{}\label{class_point3_d_a02de7d72811aef8bfc6660dd99267682} 
void {\bfseries rotate\+\_\+Z} (const double angle)
\item 
\hypertarget{class_point3_d_aa282f1dc3a75839d0581583f89eeb1ea}{}\label{class_point3_d_aa282f1dc3a75839d0581583f89eeb1ea} 
void {\bfseries rotate\+\_\+X} (const double angle)
\item 
\hypertarget{class_point3_d_af415dc7a082ad664812b02df15902217}{}\label{class_point3_d_af415dc7a082ad664812b02df15902217} 
void {\bfseries translate\+\_\+Z} (const double height)
\end{DoxyCompactItemize}
\subsection*{Friends}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_point3_d_a84c3d7aa704e44e7f6cfbfe0cb6528f7}{}\label{class_point3_d_a84c3d7aa704e44e7f6cfbfe0cb6528f7} 
void {\bfseries swap} (\hyperlink{class_point3_d}{Point3D} \&a, \hyperlink{class_point3_d}{Point3D} \&b)
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Point3d.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Point3d.\+cpp\end{DoxyCompactItemize}