### 2016-09-29 doxygen/latex/class_ray3_d.tex

\hypertarget{class_ray3_d}{}\section{Ray3D Class Reference}
\label{class_ray3_d}\index{Ray3D@{Ray3D}}

Collaboration diagram for Ray3D\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=254pt]{doxygen/latex/class_ray3_d__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_ray3_d_a0cd2092288763fc6c0fd8dccebf11527}{}\label{class_ray3_d_a0cd2092288763fc6c0fd8dccebf11527} 
{\bfseries Ray3D} (const \hyperlink{class_point3_d}{Point3D} \&origin, const \hyperlink{class_vector3_d}{Vector3D} \&direction, const double power=1.\+0, const double frequency=0.\+0, const double distance=0)
\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_ray3_d_aa303389d0d8e8569f4bdef420e850717}{}\label{class_ray3_d_aa303389d0d8e8569f4bdef420e850717} 
\hyperlink{class_point3_d}{Point3D} {\bfseries origin}
\item 
\hypertarget{class_ray3_d_a1b9fbe26d23a38f12be002d0e5990fd2}{}\label{class_ray3_d_a1b9fbe26d23a38f12be002d0e5990fd2} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries direction}
\item 
\hypertarget{class_ray3_d_adbe652ae6043219a7a6d22c3e6a454db}{}\label{class_ray3_d_adbe652ae6043219a7a6d22c3e6a454db} 
double {\bfseries power}
\item 
\hypertarget{class_ray3_d_a5be97a26ac4f733fee0b9a10ca78197a}{}\label{class_ray3_d_a5be97a26ac4f733fee0b9a10ca78197a} 
double {\bfseries frequency}
\item 
\hypertarget{class_ray3_d_a1287339cd3f7358590b2fb65dbb88505}{}\label{class_ray3_d_a1287339cd3f7358590b2fb65dbb88505} 
double {\bfseries distance}
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Ray3d.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Ray3d.\+cpp\end{DoxyCompactItemize}