### 2016-09-29 doxygen/latex/class_receiver.tex

\hypertarget{class_receiver}{}\section{Receiver Class Reference}
\label{class_receiver}\index{Receiver@{Receiver}}

Collaboration diagram for Receiver\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=195pt]{doxygen/latex/class_receiver__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_receiver_a94a089e687b5320f02b3000304ffd08c}{}\label{class_receiver_a94a089e687b5320f02b3000304ffd08c} 
{\bfseries Receiver} (const int id, const double Bandwidth, const double center\+\_\+freq, const \hyperlink{class_point3_d}{Point3D} \&center, const double boundary\+\_\+radius, const std\+::string \&savefile\+\_\+name, const std\+::string \&dopplerfile\+\_\+name)
\item 
\hypertarget{class_receiver_a6d0f579d480c7303d8138981e546ff01}{}\label{class_receiver_a6d0f579d480c7303d8138981e546ff01} 
\hyperlink{class_sphere}{Sphere} {\bfseries get\+\_\+\+Boundary} ()
\item 
\hypertarget{class_receiver_a6e97969b694ba385a495bf64f9adc16a}{}\label{class_receiver_a6e97969b694ba385a495bf64f9adc16a} 
double {\bfseries get\+\_\+sampling\+\_\+rate} ()
\item 
\hypertarget{class_receiver_a146b5c800df5f2e17eb89faa16a3929a}{}\label{class_receiver_a146b5c800df5f2e17eb89faa16a3929a} 
std\+::vector$<$ \hyperlink{class_ray3_d}{Ray3D} $>$ {\bfseries get\+\_\+frame\+\_\+data} ()
\item 
\hypertarget{class_receiver_aaa3ac7cecd33cba5a382fb0db441082e}{}\label{class_receiver_aaa3ac7cecd33cba5a382fb0db441082e} 
void {\bfseries save\+\_\+ray\+\_\+to\+Frame} (\hyperlink{class_ray3_d}{Ray3D} \&ray)
\item 
\hypertarget{class_receiver_a72ce8129ebd0720a810959eb0a47dfc8}{}\label{class_receiver_a72ce8129ebd0720a810959eb0a47dfc8} 
void {\bfseries save\+\_\+to\+\_\+file} ()
\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_receiver_aaadfb4436bd66669d03811c846985843}{}\label{class_receiver_aaadfb4436bd66669d03811c846985843} 
int {\bfseries id}
\item 
\hypertarget{class_receiver_a19225a5d1c7c5f4cdaa36b99a1184d26}{}\label{class_receiver_a19225a5d1c7c5f4cdaa36b99a1184d26} 
double {\bfseries Bandwidth}
\item 
\hypertarget{class_receiver_a33f4396a9edddf6fe298108ffa12da58}{}\label{class_receiver_a33f4396a9edddf6fe298108ffa12da58} 
double {\bfseries center\+\_\+freq}
\item 
\hypertarget{class_receiver_a7cf7efbe8c86d8f585e744cbba86c267}{}\label{class_receiver_a7cf7efbe8c86d8f585e744cbba86c267} 
double {\bfseries Resolution}
\item 
\hypertarget{class_receiver_ad9a009e1e386791e5c3bd7211d83a3f9}{}\label{class_receiver_ad9a009e1e386791e5c3bd7211d83a3f9} 
double {\bfseries boundary\+\_\+radius}
\item 
\hypertarget{class_receiver_a4432a9eafa52a2483918d2d3119cb597}{}\label{class_receiver_a4432a9eafa52a2483918d2d3119cb597} 
\hyperlink{class_point3_d}{Point3D} {\bfseries center}
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Receiver.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Receiver.\+cpp\end{DoxyCompactItemize}