### 2016-09-29 doxygen/latex/class_rib.tex

\hypertarget{class_rib}{}\section{Rib Class Reference}
\label{class_rib}\index{Rib@{Rib}}

Collaboration diagram for Rib\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=168pt]{doxygen/latex/class_rib__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_rib_a417cde052770f55f3344d785aef969ef}{}\label{class_rib_a417cde052770f55f3344d785aef969ef} 
{\bfseries Rib} (int id, \hyperlink{class_rib}{Rib} \&x, const double delta\+\_\+l)
\item 
\hypertarget{class_rib_aaaed7a2c960a33e5ac0c510ea7d8881c}{}\label{class_rib_aaaed7a2c960a33e5ac0c510ea7d8881c} 
{\bfseries Rib} (int id, const std\+::string \&filename)
\item 
\hypertarget{class_rib_ae4ea708286531b624f587f22b9416a36}{}\label{class_rib_ae4ea708286531b624f587f22b9416a36} 
std\+::vector$<$ \hyperlink{class_point3_d}{Point3D} $>$ {\bfseries get\+Rib\+Points} ()
\item 
\hypertarget{class_rib_ab6715214f4a43ed51b46a8f42d6f198b}{}\label{class_rib_ab6715214f4a43ed51b46a8f42d6f198b} 
void {\bfseries pitch} (double angle)
\item 
\hypertarget{class_rib_aafef170600416b0e6a86ccb4e8bd4e32}{}\label{class_rib_aafef170600416b0e6a86ccb4e8bd4e32} 
void {\bfseries rotate} (double angle)
\item 
\hypertarget{class_rib_a08859793a4800c7d160d0b494ad39af7}{}\label{class_rib_a08859793a4800c7d160d0b494ad39af7} 
void {\bfseries height} (const double height)
\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_rib_ac88341591298af4d11837a09b47917d5}{}\label{class_rib_ac88341591298af4d11837a09b47917d5} 
int {\bfseries id}
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Rib.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Rib.\+cpp\end{DoxyCompactItemize}