### 2016-09-29 doxygen/latex/class_rotor.tex

\hypertarget{class_rotor}{}\section{Rotor Class Reference}
\label{class_rotor}\index{Rotor@{Rotor}}

Collaboration diagram for Rotor\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=196pt]{doxygen/latex/class_rotor__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_rotor_af922b65832291cafdb53c52a545ac136}{}\label{class_rotor_af922b65832291cafdb53c52a545ac136} 
double {\bfseries get\+\_\+\+R\+PM} ()
\item 
\hypertarget{class_rotor_a25eef55d0719b86a255d3aa09f7a2688}{}\label{class_rotor_a25eef55d0719b86a255d3aa09f7a2688} 
int {\bfseries get\+\_\+num\+\_\+blades} ()
\item 
\hypertarget{class_rotor_ad0c276d01564fbdda8ca63c7eff18df6}{}\label{class_rotor_ad0c276d01564fbdda8ca63c7eff18df6} 
double {\bfseries get\+\_\+height} ()
\item 
\hypertarget{class_rotor_a53d18f1d59a1c2b71cdb8c9cab0acd4c}{}\label{class_rotor_a53d18f1d59a1c2b71cdb8c9cab0acd4c} 
double {\bfseries get\+\_\+constant\+\_\+pitch} ()
\item 
\hypertarget{class_rotor_af384431ad8eba046b8541ac5d1a3f9f9}{}\label{class_rotor_af384431ad8eba046b8541ac5d1a3f9f9} 
double {\bfseries get\+\_\+\+Rib\+\_\+count} ()
\item 
\hypertarget{class_rotor_a43485a289837aab864e28dc00109fdf6}{}\label{class_rotor_a43485a289837aab864e28dc00109fdf6} 
double {\bfseries get\+\_\+blade\+\_\+length} ()
\item 
\hypertarget{class_rotor_a31c1b1a9674d87d0b612db4989a4ae16}{}\label{class_rotor_a31c1b1a9674d87d0b612db4989a4ae16} 
std\+::vector$<$ \hyperlink{class_blade__surface}{Blade\+\_\+surface} $>$ {\bfseries get\+\_\+\+Blades} ()
\item 
\hypertarget{class_rotor_ac73f8ab43793bc45b978c6ef14239a4c}{}\label{class_rotor_ac73f8ab43793bc45b978c6ef14239a4c} 
void {\bfseries rotate} (const double angle)
\item 
\hypertarget{class_rotor_ab1ecb2c8efc6e125db1fe9121c6c47b9}{}\label{class_rotor_ab1ecb2c8efc6e125db1fe9121c6c47b9} 
bool {\bfseries hit} (const \hyperlink{class_ray3_d}{Ray3D} \&ray, double \&hit\+Distance, \hyperlink{class_vector3_d}{Vector3D} \&hit\+Normal, \hyperlink{class_point3_d}{Point3D} \&hit\+Point)
\item 
\hypertarget{class_rotor_ad525c69ea4ba137f953675839befcaa6}{}\label{class_rotor_ad525c69ea4ba137f953675839befcaa6} 
{\bfseries Rotor} (const int id, const int num\+\_\+blades, const double R\+PM, const double height, const double constant\+\_\+pitch, const double blade\+\_\+length, const int Rib\+\_\+count)
\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_rotor_abcb7f6ed3121a0734eef3763d4d20172}{}\label{class_rotor_abcb7f6ed3121a0734eef3763d4d20172} 
int {\bfseries id}
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Rotor.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Rotor.\+cpp\end{DoxyCompactItemize}