### 2016-09-29 doxygen/latex/class_sphere.tex

\hypertarget{class_sphere}{}\section{Sphere Class Reference}
\label{class_sphere}\index{Sphere@{Sphere}}

Collaboration diagram for Sphere\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=161pt]{doxygen/latex/class_sphere__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_sphere_a7b5d0be5fd60ba94573e3941cf4eb094}{}\label{class_sphere_a7b5d0be5fd60ba94573e3941cf4eb094} 
{\bfseries Sphere} (const double radius, const \hyperlink{class_point3_d}{Point3D} \&center)
\item 
\hypertarget{class_sphere_a5da44ddc57c5a68a1092b0df86dc18cd}{}\label{class_sphere_a5da44ddc57c5a68a1092b0df86dc18cd} 
bool {\bfseries hit} (const \hyperlink{class_ray3_d}{Ray3D} \&ray, double \&hit\+Distance, \hyperlink{class_vector3_d}{Vector3D} \&hit\+Normal, \hyperlink{class_point3_d}{Point3D} \&hit\+Point) const
\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_sphere_a45d6c6c870fac7c2a885ad2e226334ad}{}\label{class_sphere_a45d6c6c870fac7c2a885ad2e226334ad} 
double {\bfseries radius}
\item 
\hypertarget{class_sphere_a774cb90ed1b46592e63a112338f1cc58}{}\label{class_sphere_a774cb90ed1b46592e63a112338f1cc58} 
\hyperlink{class_point3_d}{Point3D} {\bfseries center}
\item 
\hypertarget{class_sphere_a50d3abf7b35058bae9bcbba998a82943}{}\label{class_sphere_a50d3abf7b35058bae9bcbba998a82943} 
int {\bfseries num\+\_\+points}
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Sphere.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Sphere.\+cpp\end{DoxyCompactItemize}