### 2016-09-29 doxygen/latex/class_transmitter.tex

\hypertarget{class_transmitter}{}\section{Transmitter Class Reference}
\label{class_transmitter}\index{Transmitter@{Transmitter}}

Collaboration diagram for Transmitter\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=177pt]{doxygen/latex/class_transmitter__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_transmitter_a91c51bc1d7c42db81fb390753de5c0ea}{}\label{class_transmitter_a91c51bc1d7c42db81fb390753de5c0ea} 
{\bfseries Transmitter} (const int id, const double frequency, const double power, const \hyperlink{class_point3_d}{Point3D} \&center, const double l)
\item 
\hypertarget{class_transmitter_aa26f745526cba27666812ca7c05b9df7}{}\label{class_transmitter_aa26f745526cba27666812ca7c05b9df7} 
void {\bfseries set\+Power} (const double power)
\item 
\hypertarget{class_transmitter_a72eb7b0c29d6d5e3429f9920dcdf86c9}{}\label{class_transmitter_a72eb7b0c29d6d5e3429f9920dcdf86c9} 
double {\bfseries get\+Power} () const
\item 
\hypertarget{class_transmitter_a8a2727ad1e4b48ec9f64c0285bb5bd04}{}\label{class_transmitter_a8a2727ad1e4b48ec9f64c0285bb5bd04} 
\hyperlink{class_ray3_d}{Ray3D} {\bfseries make\+Ray} ()
\item 
\hypertarget{class_transmitter_a7c51cee738e95fb6f2af955986873b7c}{}\label{class_transmitter_a7c51cee738e95fb6f2af955986873b7c} 
\hyperlink{class_ray3_d}{Ray3D} {\bfseries make\+Ray\+\_\+disk} (const double height)
\item 
\hypertarget{class_transmitter_a5c6bdf0d0dd931f9f4282a7d585f2320}{}\label{class_transmitter_a5c6bdf0d0dd931f9f4282a7d585f2320} 
\hyperlink{class_ray3_d}{Ray3D} {\bfseries make\+Ray} (const \hyperlink{class_vector3_d}{Vector3D} \&ray\+Direction)
\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_transmitter_a3e57cbedcf7ff80c3506c47d43119f5d}{}\label{class_transmitter_a3e57cbedcf7ff80c3506c47d43119f5d} 
int {\bfseries id}
\item 
\hypertarget{class_transmitter_ade036fee92a7f2aad8ecf206f89b8c34}{}\label{class_transmitter_ade036fee92a7f2aad8ecf206f89b8c34} 
double {\bfseries frequency}
\item 
\hypertarget{class_transmitter_a61c6aa4371c98b49d2d9277fd4447f34}{}\label{class_transmitter_a61c6aa4371c98b49d2d9277fd4447f34} 
\hyperlink{class_point3_d}{Point3D} {\bfseries center}
\item 
\hypertarget{class_transmitter_ac339f9fd3011afdbdc729b99b371e617}{}\label{class_transmitter_ac339f9fd3011afdbdc729b99b371e617} 
double {\bfseries power}
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Transmitter.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Transmitter.\+cpp\end{DoxyCompactItemize}