### 2016-09-29 doxygen/latex/class_vector3_d.tex

\hypertarget{class_vector3_d}{}\section{Vector3D Class Reference}
\label{class_vector3_d}\index{Vector3D@{Vector3D}}

Collaboration diagram for Vector3D\+:
\nopagebreak
\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[width=155pt]{doxygen/latex/class_vector3_d__coll__graph}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hypertarget{class_vector3_d_a9c58f21ac5a280f62067d7f26466ba47}{}\label{class_vector3_d_a9c58f21ac5a280f62067d7f26466ba47} 
{\bfseries Vector3D} (const double x, const double y, const double z)
\item 
\hypertarget{class_vector3_d_a5fc1291699c792e6acd5ea453c47f3f4}{}\label{class_vector3_d_a5fc1291699c792e6acd5ea453c47f3f4} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries operator-\/} (void) const
\item 
\hypertarget{class_vector3_d_afa1d5717e088826b6dcc0cb35ec836f7}{}\label{class_vector3_d_afa1d5717e088826b6dcc0cb35ec836f7} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries operator$\ast$} (const double a) const
\item 
\hypertarget{class_vector3_d_a86a2dec32df2740e92e06b651d693c95}{}\label{class_vector3_d_a86a2dec32df2740e92e06b651d693c95} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries operator/} (const double a) const
\item 
\hypertarget{class_vector3_d_af75a42c25ca6e999f3e7a83d4de105c6}{}\label{class_vector3_d_af75a42c25ca6e999f3e7a83d4de105c6} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries operator+} (const \hyperlink{class_vector3_d}{Vector3D} \&v) const
\item 
\hypertarget{class_vector3_d_acd7927ced75a36b459200a791e51acad}{}\label{class_vector3_d_acd7927ced75a36b459200a791e51acad} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries operator-\/} (const \hyperlink{class_vector3_d}{Vector3D} \&v) const
\item 
\hypertarget{class_vector3_d_a307962a6472b58ac85af69ad74c4fd6f}{}\label{class_vector3_d_a307962a6472b58ac85af69ad74c4fd6f} 
double {\bfseries operator$\ast$} (const \hyperlink{class_vector3_d}{Vector3D} \&b) const
\item 
\hypertarget{class_vector3_d_ac6d8d994b5c80eaaed1f033fd7baa627}{}\label{class_vector3_d_ac6d8d994b5c80eaaed1f033fd7baa627} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries operator$^\wedge$} (const \hyperlink{class_vector3_d}{Vector3D} \&v) const
\item 
\hypertarget{class_vector3_d_ac86d9a556616f4c51fecbaf7d8658ce8}{}\label{class_vector3_d_ac86d9a556616f4c51fecbaf7d8658ce8} 
double {\bfseries x} () const
\item 
\hypertarget{class_vector3_d_a7146d35db2c1bf7606c8290282a423ec}{}\label{class_vector3_d_a7146d35db2c1bf7606c8290282a423ec} 
double {\bfseries y} () const
\item 
\hypertarget{class_vector3_d_af4ae427e82a23af845f804106daed509}{}\label{class_vector3_d_af4ae427e82a23af845f804106daed509} 
double {\bfseries z} () const
\item 
\hypertarget{class_vector3_d_af396bfa140d140d818a97100f7d6a532}{}\label{class_vector3_d_af396bfa140d140d818a97100f7d6a532} 
double {\bfseries dot\+Product} (const \hyperlink{class_vector3_d}{Vector3D} \&v) const
\item 
\hypertarget{class_vector3_d_a15b417a5abd07b8a7403bdb0ac5f95ec}{}\label{class_vector3_d_a15b417a5abd07b8a7403bdb0ac5f95ec} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries cross\+Product} (const \hyperlink{class_vector3_d}{Vector3D} \&v) const
\item 
\hypertarget{class_vector3_d_a4ddb90e85e10f0f32475725cb966a5ce}{}\label{class_vector3_d_a4ddb90e85e10f0f32475725cb966a5ce} 
double {\bfseries length} () const
\item 
\hypertarget{class_vector3_d_aac87aef508e65eb8ada23346b5e552c3}{}\label{class_vector3_d_aac87aef508e65eb8ada23346b5e552c3} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries normalized} () const
\item 
\hypertarget{class_vector3_d_a1698a0e33c975b78f0076b51331e6e94}{}\label{class_vector3_d_a1698a0e33c975b78f0076b51331e6e94} 
\hyperlink{class_vector3_d}{Vector3D} {\bfseries rotated\+AboutZ} (const double angle) const
\item 
\hypertarget{class_vector3_d_a5df051b736bdb50b63995442b0816938}{}\label{class_vector3_d_a5df051b736bdb50b63995442b0816938} 
bool {\bfseries is\+Normal} () const
\end{DoxyCompactItemize}

The documentation for this class was generated from the following files\+:\begin{DoxyCompactItemize}
\item 
Raytracer3\+D/\+Raytracer3\+D/Vector3d.\+hpp\item 
Raytracer3\+D/\+Raytracer3\+D/Vector3d.\+cpp\end{DoxyCompactItemize}